Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-06-28 GWLDCI.tex
exposition in paper 1 and looked over 4.5

Therefore, any basis of $W$ can be written as $B \sqcup U$ for some $B$ and $U$ of the form indicated described above. Applying the same argument to $W'$

@@

If not, there would be a cyclic flat $F_j$ of $M$. Without loss of generality, write $F_j \in \mathcal{F} \setminus \mathcal{F}'$.  We may write $F_j = V(P_j)^c$, with $(P_j, V(P_j))$ a maximal exact subdiagram in $W$. By Lemma \ref{lem facts about WLD matroids} (5), we may write $F_j= F(Q)$ as a propagator flat. Therefore, Lemma \ref{maxexactcomplementrank}, $\rk(F_j) = |Q|$. By Lemma \ref{contractsubdiaglem}, $M/F(Q) = F/F_j = M(Q^c, V(Q^c))$, which is uniform, by construction of $F_j$. Theorem \ref{exactuniformthm} $(Q^c, V(Q^c))$ is exact. Since this exact subdiagram is not in $\mathcal{R}'$ by construction of $F_j$, it is not a maximal exact subdiagram of $W'$. That is, there is a maximal exact subdiagram containing it: $(Q^c, V(Q^c)) \subsetneq (P_j', V(P_j'))$. Thus there is a $F_j' \in \mathcal{R}'$ such that $F_j' \subsetneq F_j$. We show that $F_j$ is not a cyclic flat.

@@

\begin{proof}[Proposed alternative proof]
One direction has been proved in \cite[Theorem 1.18]{wilsonloop}, but we give a different proof here to be consistent with the method of this paper.

@@

In this section we have characterized the correspondence between Wilson loop diagrams and positroids, showing that Wilson Loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). Corollary \ref{number of equiv diagrams} enumerates the fibres of the map from Wilson Loop diagrams to positroids.  The obvious next question would be to enumerate the number of equivlence classes.  We do not obtain a formula for this, but do obtain an alternate characterization of the equvalence classes which is the subject of the next section.

@@

The connection between admissible Wilson loop diagrams and polytope dissections suggests a connection to associahedra. First we need some background. \sanote{expand}

@@

Note that Wilson loop diagrams which are weakly admissible but not admissible are necessarily exact, and hence have $\tau$ graphs which are triangulations, i.e. dimension 0 faces (vertices) in $A_n$.  All of these are equivalent (as Wilson loop diagrams) in the sense that they are only retriangulations away from each other; meanwhile, as noted above, all vertices of $A_n$ are parallel to each other.  It will be convenient in what follows to include these diagrams, but the results are easily modified to remove them if needed: Propositions~\ref{prop easy way} and \ref{prop hard way} would be identical if we replaced ``weakly admissible'' by ``admissible'', while Corollary~\ref{cor:count inequiv diagrams} could be changed to consider only admissible diagrams by subtracting 1 from the number of non-parallel faces in order to account for the missing dimension 0 faces.

@@

We also need the following notation: given a polygon dissection $D$ of an $n$-gon, let $V_{>3}(D)$ be the set of vertices of the polygon which are on at least one non-triangle face of $D$.  (Note that $D$ may have many non-triangle faces and so $V_{>3}(D)$ will be the union of the vertices of each non-triangle face.) We also write $[x_i,x_j]$ for the set of vertices from $x_i$ to $x_j$ including the endpoints, and $(x_i,x_j)$ for the same set but excluding the endpoints.

@@

  As observed in the proof of Proposition~\ref{prop easy way}, the support of any non-zero vector parallel to an edge is on the indices corresponding to the four vertices bounding the quadrilateral of the dissection associated with the edge.  Thus if $V_{>3}(\tau(W)) \neq V_{>3}(\tau(X))$ (notation as defined before Lemma~\ref{lem good quads}) then the span of the vectors parallel to their edges has different support and so they cannot be parallel.  So we can assume that $V_{>3}(\tau(W)) = V_{>3}(\tau(X))$.

@@

Propositions~\ref{prop easy way} and \ref{prop hard way} together imply:
\begin{cor}\label{cor:count inequiv diagrams}
  The number of inequivalent weakly admissible Wilson loop diagrams on $n$ is the number of non-parallel faces in the associahedron $A_n$.
\end{cor}

>>>>

Therefore, any basis of $W$ can be written as $B \sqcup U$ for some $B$ and $U$ of the form indicated described above. We may apply the same arguments to $W'$.

@@

If not, there would be a cyclic flat $F_j$ of $M$ in one set but not the other. Without loss of generality, write $F_j \in \mathcal{F} \setminus \mathcal{F}'$.  We may write $F_j = V(P_j)^c$, with $(P_j, V(P_j))$ a maximal exact subdiagram in $W$. By Lemma \ref{lem facts about WLD matroids} (5), we may write $F_j= F(Q)$ as a propagator flat. Therefore, Lemma \ref{maxexactcomplementrank}, $\rk(F_j) = |Q|$. By Lemma \ref{contractsubdiaglem}, $M/F(Q) = F/F_j = M(Q^c, V(Q^c))$, which is uniform, by construction of $F_j$. Theorem \ref{exactuniformthm} $(Q^c, V(Q^c))$ is exact. Since this exact subdiagram is not in $\mathcal{R}'$ by construction of $F_j$, it is not a maximal exact subdiagram of $W'$. That is, there is a maximal exact subdiagram containing it: $(Q^c, V(Q^c)) \subsetneq (P_j', V(P_j'))$. Thus there is a $F_j' \in \mathcal{R}'$ such that $F_j' \subsetneq F_j$. We show that $F_j$ is not a cyclic flat.

@@

\begin{proof}[Proposed alternative proof] \sanote{Looks good to me}
One direction has been proved in \cite[Theorem 1.18]{wilsonloop}, but we give a different proof here to be consistent with the method of this paper.

@@

In this section we have characterized the correspondence between Wilson loop diagrams and positroids, showing that Wilson Loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). Corollary \ref{number of equiv diagrams} enumerates the fibres of the map from Wilson Loop diagrams to positroids.  The obvious next question would be to enumerate the number of equivlence classes.

We know from example, for instance \cite{casestudy} that, even after taking the dimension of the positroid cell into consideration\footnote{it is proven in \cite{GWLDII} and \cite{Marcott} that Wilson loop diagrams with $k$ propagators map to $3k$ dimensional positroid cells}, many positroid cells do not appear as admissible Wilson loop diagrams with a fixed number of propagators and vertices. That is, the map from admissible Wilson loop diagrams to positroid cells of the correct dimension in the correct Grassmannian neither injective nor surjective. We do not obtain a formula for the number of cells, but do obtain an alternate characterization of the equvalence classes which is the subject of the next section.

@@

The correspondence between (weakly) admissible Wilson loop diagrams and polygon dissections shown in section \ref{sec: polygon partitions} suggests a connection to associahedra. First we need some background. There is a realization of associahedra that is charactertized by polygon partitions. In this section, we show that the number of equivalence classes of Wilson loop diagrams correspond to the number of non-parallel faces of the appropriate associahedron (Theorem \ref{thm:count }). While this section does not give an explicit formula for the number of these faces, we hope that this linkage with a combinatorially well studied object will shed light on the issue in the future.

@@

Note that Wilson loop diagrams which are weakly admissible but not admissible are necessarily exact, and hence have $\tau$ graphs which are triangulations, i.e. dimension 0 faces (vertices) in $A_n$.  All of these are equivalent (as Wilson loop diagrams) in the sense that they are only retriangulations away from each other; meanwhile, as noted above, all vertices of $A_n$ are parallel to each other.  It will be convenient in what follows to include these diagrams, but the results are easily modified to remove them if needed: Propositions~\ref{prop easy way} and \ref{prop hard way} would be identical if we replaced ``weakly admissible'' by ``admissible'', while Theorem~\ref{thm:count inequiv diagrams} could be changed to consider only admissible diagrams by subtracting 1 from the number of non-parallel faces in order to account for the missing dimension 0 faces.

@@

We also need the following notation: given a polygon dissection $D$ of an $n$-gon, let $V_{>3}(D)$ be the set of vertices of the polygon which are on at least one non-triangle face of $D$.  (Note that $D$ may have many non-triangle faces and so $V_{>3}(D)$ will be the union of the vertices of each non-triangle face.) We also write $[x_i,x_j]$ for the set of vertices from $x_i$ to $x_j$ including the endpoints, and $(x_i,x_j)$ for the same set but excluding the endpoints. \sanote{make a dfn?}

@@

As observed in the proof of Proposition~\ref{prop easy way}, \st{the support of any non-zero vector parallel to an edge is on the indices corresponding to the four vertices bounding the quadrilateral of the dissection associated with the edge.}{\color{violet}any non-zero vector in $\R^n$ paralell to an edge of $A_n$ has non-zero entries only in the coordinates corresponding to the four vertices bounding the quadrilateral of the dissection associated with the edge.}  Thus if $V_{>3}(\tau(W)) \neq V_{>3}(\tau(X))$ \hlfix{(notation as defined before Lemma {\ref{lem good quads}})}{refer to dfn?} then the span of the vectors parallel to their edges has different support and so they cannot be parallel.  So we can assume that $V_{>3}(\tau(W)) = V_{>3}(\tau(X))$.

@@

Propositions~\ref{prop easy way} and \ref{prop hard way} together imply the main result of this section:
\begin{thm}\label{thm:count inequiv diagrams}
  The number of inequivalent weakly admissible Wilson loop diagrams on $n$ is the number of non-parallel faces in the associahedron $A_n$.
\end{thm} \sanote{this is the main result of the section}
